Fixing Vulnerabilities Hinders Maintainability

### 2019-02-01 paper/ICPC19/icpc19.tex
discussion up

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/bch_report.png}
 	\caption{Maintainability report of OpenSSL CVE-2014-3506 vulnerability
refactoring for the guideline \emph{Write Simple Units of Code} provided by
\emph{Better Code Hub}. This version of OpenSSL does not comply with the
guideline in the example since the bars do not reach the threshold points. This
example only complies with $\sfrac{1}{10}$ guidelines (\emph{Write Clean Code}).}
	\label{fig:bchrep}
\end{figure}

@@

\subsection{Statistical Validation}\label{sec:statsval}
%
To validate the maintainability differences in different groups of commits
(e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank
test with the significance level $\alpha = 0.05$ \cite{10.2307/3001968} . In
other words, we test the null hypothesis that the maintainability difference
between pairs of versions $v_{v-1}$, $v_v$ (i.e., before and after a
security-commit) follows a symmetric distribution around $0$. However, when the
difference between groups is zero, the observations are discarded. Thus, we use
a more robust approach of the wilcoxon test, the Wilcoxon test by Pratt which
provides an alternative that incorporates the cases where the maintainability
difference is zero \cite{10.2307/2282543}. To understand the effect-size, as
advocated by the Common-language effect sizes \cite{graw:1992}, we compute the
mean difference, the median of the difference, and the percentage of cases that
reduce maintainability.

%
\section{Results}\label{sec:results}

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.55\textwidth]{figures/language.pdf}
 	\caption{Maintainability difference by language after security refactorings}
	\label{fig:lang}
\end{figure}

In this study, we also evaluated the impact of each programming language on the 
maintainability of open-source software, Figure \ref{fig:lang}. Although the
dataset contains security refactorings for more than $13$ languages, we only
present the results for \emph{Java}, \emph{C}, \emph{Ruby} and \emph{PHP} because
one of the requirements of the Wilcoxon signed-rank test is that the input sample
has to have more than $20$ measurements. The other languages were integrated in the
\emph{Others} set. The results of the Wilcoxon signed-rank test by pattern returned 
statistical evidence which supports that the maintainability of open-source software 
decreases for security refactorings in 3 different languages: \emph{C} 
(p = $5.72$x$10^{-9}$), Ruby (p = 0.057) and PHP (p = $3.54$x$10^{-6}$). The results 
also provide statistical significance for when security refactorings are performed
in \emph{Java} (p = $3.61$x$10^{-2}$). \emph{Python} and the other languages 
that integrate the \emph{Others} set contain more cases where the maintainability
increases. However, the results obtained for these patterns are not statistically 
significant.

@@

Projects with Broken Authentication flaws are the most affected ones by security refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any change. The results also show that fixing memory leaks affects the maintainability of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected. 35\% (28) are affected positively. Refactoring vulnerabilities that allow denial-of-service attacks (e.g., CVE-2014-3506) has a negative impact of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of the cases exhibit maintainability improvement while 22.5\% (9) show that overall maintainability does not suffer any change. Miscellaneous refactorings pattern integrates several types of security flaws whose impact is of 43.1\% (84) negative in the maintainability of open-source sofware. 31.8\% (62) of the cases do not exhibit any change in maintainability and 25.1\% (49) of cases have a positive impact on the maintainability of open-source software.

@@

40.2\% of the security refactoring have a negative impact on the maintainability of open-source software, Figure \ref{fig:secvsreg}.

% find work about the struggle to make systems secure

We believe developers should be able to create secure code without affecting 
the maintainability of their projects.

\begin{itemize}
	\item \textbf{Security Frameworks.} The most popular languagues (e.g., Java and c\#) 
used in the development of enterprise applications provide stable
frameworks to implement the security requirements and easily prevent some of the
OWASP TOP 10 vulnerabilities.
	\item \textbf{Documentation}
	\item \textbf{Programming Languages}
	\item \textbf{Proper authorization mechanisms}
	\item \textbf{Remove dead and duplicated code}
	\item \textbf{Configurations in files}
	\item \textbf{Framework manages the session}

@@

% #1 Security frameworks are your new best friend
% The most popular software stacks (e.g.: Java and C#) used in the development of enterprise applications provide very good and stable frameworks to implement the security requirements and easily prevent some of the OWASP TOP 10 vulnerabilities. For instance, through a very simple annotation such as @EnableWebSecurity, Spring Security automatically and out of the box, implements mitigations against some of the most common web application vulnerabilities (e.g.: CSRF, XSS, etc.). The essential HTTP headers to prevent certain types of vulnerabilities are added to the requests and validated in the responses (e.g.: same-origin policy for the synchronization token to prevent CSRF).
%
% Although these frameworks may require some time to understand and use (the learning curve might be quite steep, sometimes), they have been improving quite a lot in the past few years, facilitating the implementation and maintenance of the security requirements.
%
% Nevertheless, do not blindly trust all the defaults of the frameworks and libraries your application depends on. More than often frameworks rely on weak default encryption algorithms such as DES (encryption standard deprecated since 1998). Thus, always check all the frameworks and libraries defaults, before releasing the application.
%
% Besides, some of the least maintained frameworks contain vulnerabilities that never get to be reported. This is because some bugs (this is an interesting discussion about a systemd null pointer dereferencing bug) are not even considered security vulnerabilities in the first place. Notice that categorisation of bugs as security vulnerabilities, is based on the ability of the development team to evaluate the impact of that bug in the security of the application. That’s why updating the libraries and frameworks of your application to their latest version helps improving its overall security. So, stay up-to-date.
%
% #2 Deny first, authorize after
% Broken access control is one of the OWASP TOP 10 vulnerabilities. This is because most applications fail to implement proper authorization mechanisms. Authorization design and implementation is probably one of the most complex security features that exists. Making it right and scalable is challenging and requires the effort of all the team members. Nevertheless, adoption of certain mechanisms and design strategies may facilitate authorization maintenance in the long run:
%
% 1. Authorise only afterwards: access must always be denied in the first place, i.e., access must only be granted if the user is actually authorized to perform a certain action. So, access control policies must be checked before processing the request and not when producing the response.
%
% 2. Use and abuse of filters: HTTP filters can be used to filter out all the requests before processing them (e.g.: if the user doing the request is authorized to perform the designated action, then the request can be processed). Certain frameworks (e.g.: Spring Security) even provide a very elegant way of binding the access control policies to certain system functions by means of code annotations in the methods that expose the desired functionalities. In combination with the HTTP filters these annotations provide a very clean and easy to change implementation of the access control mechanisms.
%
% 3. Always document what you’re doing: either using role-based access control or a more fine-grained access control model, the functionalities enabled in each roles must be clearly defined and documented. Although, the code should speak for itself, the lack of documentation brings risks for code portability and transferability, specially when considering access control management.
%
% #3 Make the IT infrastructure your number one software project
% A securely developed application is defenseless against an attacker with administrative privileges in the production environment. For instance, the WannaCry ransomware attack was based on an OS level exploit that was fairly easy to protect, just by upgrading the OS version. Nonetheless, it was very problematic because a lot of IT departments tend to disregard the importance and impact of OS security updates.
%
% In general, the longer you wait to upgrade, the harder it gets and eventually a system can get ‘un-upgradable’, therefore, unmaintainable.
% When deploying the application, it is necessary to make sure that all 3rd party components that are part of the production environment can be upgraded or recreated from backup (in case of ransomware) at any point in time. This includes databases and web servers, deployment frameworks, the identity manager, log manager application and alerting system, the operating system, etc.
%
% The ‘upgrading is hard’ problem is similar to the ‘releasing is hard’ problem and should be solved the same way: by doing it often.
% The way to do this is to treat the environment in which your application runs as part of the application and not something separate. As such it should be treated as a software project in itself.
%
% #4 Validations should be done in the backend
% More and more web applications follow a service oriented architecture model, where the backend provides a RESTful API and the front-end is purely implemented in JavaScript (with HTML). Some modern JavaScript frameworks provide out of the box escaping and sanitization functionalities (e.g.: Angular 2 or above). Some developers are led to mistakenly believe that this free-feature, together with type checking the REST parameters (done in the corresponding controller, for instance) is enough to prevent introduction of malicious input in the application. But unfortunately, it is not. As much as the web and security frameworks try to enable security mitigations by default, a lot of effort is still needed from the development team. Therefore, it’s very important to validate all the user input in the backend and filter-out unwanted/invalid inputs.
%management of the web application.
%
% Notice that the fact that in modern frameworks crypto primitives can be defined in a configuration file, makes it much easier and less time consuming, to, for instance, update those primitives when a new attack is discovered. Plus, if the default session manager has a known vulnerability, it is easier to patch the application by updating the library, then doing quick (and perhaps dirty) fixes in custom implementations, that can lead to even more exploits.
%

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

\textbf{\textit{Broken Authentication \& Session Management, Denial-of-Service attacks and Memory Leaks and Miscellaneous patterns are more likely to affect open-source software maintainability.}}

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

>>>>

\Sof{Models} Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions.

@@

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/bch_report.png}
 	\caption{Maintainability report of OpenSSL CVE-2014-3506 vulnerability
refactoring for the guideline \emph{Write Simple Units of Code} provided by
\emph{Better Code Hub}. This version of OpenSSL does not comply with the
guideline in the example since the bars do not reach the threshold points. This
example only complies with $\frac{1}{10}$ guidelines (\emph{Write Clean Code}).}
	\label{fig:bchrep}
\end{figure}

@@

\subsection{Statistical Validation}\label{sec:statsval}
%
To validate the maintainability differences in different groups of commits
(e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank
test with the significance level $\alpha = 0.05$ \cite{10.2307/3001968} . In
other words, we test the null hypothesis that the maintainability difference
between pairs of versions $v_{v-1}$, $v_v$ (i.e., before and after a
security-commit) follows a symmetric distribution around $0$. However, when the
difference between groups is zero, the observations are discarded. Thus, we use
a more robust approach of the wilcoxon test, the Wilcoxon test by Pratt which
provides an alternative that incorporates the cases where the maintainability
difference is zero \cite{10.2307/2282543}. To understand the effect-size, as
advocated by the Common-language effect sizes \cite{graw:1992}, we compute the
mean difference, the median of the difference, and the percentage of cases that
reduce maintainability.
%
\section{Results}\label{sec:results}

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.55\textwidth]{figures/language.pdf}
 	\caption{Maintainability difference by language after security refactorings}
	\label{fig:lang_main}
\end{figure}

In this study, we also evaluated the impact of each programming language on the 
maintainability of open-source software, Figure \ref{fig:lang_main}. Although the
dataset contains security refactorings for more than $13$ languages, we only
present the results for \emph{Java}, \emph{C}, \emph{Ruby} and \emph{PHP} because
one of the requirements of the Wilcoxon signed-rank test is that the input sample
has to have more than $20$ measurements. The other languages were integrated in the
\emph{Others} set. The results of the Wilcoxon signed-rank test by pattern returned 
statistical evidence which supports that the maintainability of open-source software 
decreases for security refactorings in 3 different languages: \emph{C} 
(p = $5.72$x$10^{-9}$), Ruby (p = 0.057) and PHP (p = $3.54$x$10^{-6}$). The results 
also provide statistical significance for when security refactorings are performed
in \emph{Java} (p = $3.61$x$10^{-2}$). \emph{Python} and the other languages 
that integrate the \emph{Others} set contain more cases where the maintainability
increases. However, the results obtained for these patterns are not statistically 
significant.

@@

Projects with Broken Authentication flaws are the most affected ones by security 
refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability 
while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any 
change. The results also show that fixing memory leaks affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected. 
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities 
that allow denial-of-service attacks (e.g., CVE-2014-3506) has a negative impact 
of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of 
the cases exhibit maintainability improvement while 22.5\% (9) show that overall 
maintainability does not suffer any change. Miscellaneous refactorings pattern 
integrates several types of security flaws whose impact is of 43.1\% (84) negative 
in the maintainability of open-source sofware. 31.8\% (62) of the cases do not 
exhibit any change in maintainability and 25.1\% (49) of cases have a positive 
impact on the maintainability of open-source software.

@@

40.2\% of the security refactoring have a negative impact on the maintainability 
of open-source software, Figure \ref{fig:secvsreg}. This raises a new tradeoff when
developers need to fix vulnerabilities on their projects because they might be affecting 
the software maintainability or even introducing new vulnerabilities. This study exhibits
proof that developers may have to reduce maintainability for the sake of safety. We understand 
that developers should be able to produce secure code without affecting the maintainability of 
their projects. But there are still some concerns that need to be addressed: 
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure. Unfortunately, there are several 
	vulnerabilities that stay undisclosed. Thus, developers should not trust them blindly and 
	check defaults.
	\item \textbf{Documentation} of software should provide developers with the best practices to implement safe and maintainable software. Unmaintainable code is difficult to test, analyze and re-use.
	\item \textbf{Proper Authorization Mechanisms} should be designed and implemented. This is
	one of the most complex security features to implement. Some mechanisms might assist authorization maintenance in long run, such as the combination of \texttt{HTTP} filters with \texttt{Spring Security} annotations.
	\item \textbf{Deprecated code} should be avoided by developers. They should integrate tools such as BCH on their projects CI/CD pipelines to help them be more aware of the maintainability of their code and issues (e.g., dead and duplicated code). 
	\item\textbf{Programming Languages} should provide new mechanisms to easily implement security patterns without endangering maintainability.

@@

We would expect that the negative impact of programming languages on maintainability 
would be more intense due to the buzz of the last years surrounding the idea that 
poor design of programming languages and lack of using the best practices lead to 
more buggy and vulnerable code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}. 
However, Figure \ref{fig:lang_main} shows that only \emph{C} has a negative impact of more than 50\% on 
software maintainability. We suspect that these values are the result of projects 
contributions policies. 9 out of 10 projects --- the equivalent to 31.8\% of the commits --- 
of the top-10 of dataset projects with more contributors have restrict contribution 
policies regarding code standards. Some examples are the \texttt{laravel/framework}, 
\texttt{php/php-src}, \texttt{openssl/openssl} and 
\texttt{cakephp/cakephp}. For instance, \texttt{laravel/framework} requires the 
\texttt{PSR-2 Standard}\footnote{PSR-2 Standard available at 
https://www.php-fig.org/psr/psr-2/ (Accessed on \today{}}).

@@

\textbf{\textit{Broken Authentication \& Session Management, Denial-of-Service attacks and Memory Leaks and Miscellaneous patterns are more likely to affect open-source software maintainability.}} Although there are
a few refactorings that harm software maintainability in the \emph{Cross-Site Scripting} and \emph{Using 
Components with Know Vulnerabilities}, overall maintainability does not suffer any changes for these patterns.
For the remaining patterns, \emph{Injection} and \emph{Cross-Site Request Forgery}, results are not statistically
significant.

The impact of a refactoring depends on its complexity, i.e., if the refactoring adds complexity to the codebase it is probably affecting the software maintainability. One evidence of that is the fact of \emph{Cross-Site Scripting} and \emph{Using Components with Know Vulnerabilities} refactorings endure more cases without any impact in the maintainability. Typically, these type of vulnerabilities are one line fixes --- adding an \texttt{escape} function or changing the version of the component, respectively. Whereas, Broken Authentication and vulnerabilities such as the one presented in Section \ref{sec:motivation} responsible by DoS attacks are more difficult to fix (several code lines deleted and added). Another example is the Apple authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872 details available at https://support.apple.com/en-us/HT208315 (Accessed on \today{}}, where any user could login as root with an empty password. Patrick Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on \today{}} and conclude that the flaw was due to the introduction of high cyclomatic complexity in a method used to verify the password. In sum, \emph{Cross-Site Scripting} and \emph{Using Components with Known Vulnerabilities} should take part of the developers preocupations along with the other patterns that explicitly hinder maintainability because despite their low complexity these are patterns that still produce a considerable negative impact on the maintainability. Real efforts have been made in creating and incorporating coding standards and best practices in software production. However, security is still far from perfection and fixing vulnerabilities might hinder software maintainability. Thus, it is important to provide tool support to developers to help them applying these patterns without harming software maintainability.

@@

@@
